---
title: "R Notebook"
output: html_notebook
---

# Notes  

Some highlights:

*  Using `httr::` and `readxl::` packages to pull down data from data.world.
*  Using the `zoo::as.yearmon()` function to convert into a date format.  
*  Experimenting with some small multiples techniques (inspired by kjhealy's COVID-19 visualzation: https://github.com/kjhealy/covid).  
*  Mapping the data to show divergent story from most recent data, which 

Further development:  
*  Creating `year_month` using an intermediate step, but could have used `stringr::` and `lubridate::` for more direct approach.  
*  Mapping (and the mapped story) is preliminary.  
*  Tried to implement `RColorBrewer::` in the map, but had some challenges and moved on.  

# Setup

```{r setup, include=TRUE}
library(tidyverse)
library(httr)
library(readxl)
library(zoo)
library(magrittr)
library(viridis)
#library(lubridate)
#library(RColorBrewer)
```

```{r map.theme, include=TRUE}
map.theme <- theme(text = element_text(color = "#FFFFFF"), 
                   panel.background = element_rect(fill = "#444444"),
                   plot.background = element_rect(fill = "#444444"),
                   panel.grid = element_blank(),
                   plot.title = element_text(size = 50),
                   plot.subtitle = element_text(size = 25),
                   plot.caption = element_text(size = 20), 
                   axis.text = element_blank(),
                   axis.title = element_blank(),
                   axis.ticks = element_blank(),
                   legend.position = "bottom", 
                   legend.key.size = unit(1, "cm"),
                   legend.background = element_rect(fill="#444444"),
                   legend.title = element_blank())
```

```{r read_data, include=TRUE}
GET("https://query.data.world/s/h5cw7k322n3kptttmta7q5hirzh5fa", write_disk(tf <- tempfile(fileext = ".xlsx")))
df <- read_excel(tf)
```

```{r rename_cols, include=TRUE}
df %<>% 
  slice(-1) %>% 
  rename(category = `Table 1`,
         country = `...2`,
         year_month = `...3`,
         index_2015_100 = `...4`)
```

```{r glimpse, include=TRUE}
df %>% 
  glimpse()
```

```{r refactor_mutate, include=TRUE}
(df %<>% 
  #stringr would be cleaner, but this works
  separate(year_month, into = c('year', 'month'), sep = 'M') %>% 
  mutate(year_month = paste(year, month, sep = '-'),
         index_2015_100 = as.numeric(index_2015_100)) %>% 
  mutate(year_month = as.yearmon(year_month)) %>% 
  select(1, 2, 5, 6))
```

```{r summary, include=TRUE}
df %>% 
  summary()
```

## Time series visualization  

Plotting some simple time series stories from the data.

```{r eu_viz, include=TRUE}
df %>% 
  filter(country == 'EU27_2020 - European Union - 27 countries (from 2020)',
         year_month >= 'Jan 2010') %>% 
  ggplot(.) + 
  geom_line(aes(x=year_month,
                y=index_2015_100,
                group=category,
                color=category),
            size=1.1) + 
  ggtitle('Book prices vs total consumer prices in the European Union') + 
  labs(x='',
       y='HCPI, 2015 = 100\n',
       color='') + 
  theme_minimal() + 
  theme(legend.position = 'top')
```


```{r denmark_viz, include=TRUE}
df %>% 
  filter(country == 'Denmark',
         year_month >= 'Jan 2010') %>% 
  ggplot(.) + 
  geom_line(aes(x=year_month,
                y=index_2015_100,
                group=category,
                color=category),
            size=1.1) + 
  ggtitle('Book prices vs total consumer prices in the Denmark') + 
  labs(x='',
       y='HCPI, 2015 = 100\n',
       color='') + 
  theme_minimal() + 
  theme(legend.position = 'top')
```

## Small multiples  

This code borrows heavily from https://github.com/kjhealy/covid.

```{r country_panel, include=TRUE}
#create dataframe for country names to appear in small multiples within the
#plot instead of in strip text
country_panel <- df %>%
  filter(!country %in% 
           c('United States', 
             'EU27_2020 - European Union - 27 countries (from 2020)')) %>% 
  group_by(country) %>%
  filter(year_month == 'Jan 2010') %>% 
  filter(row_number() == 1) %>% 
  ungroup()
```

```{r plot_small_multiples, fig.height=12, fig.width=10}
df  %>%
  filter(!country %in% 
           c('United States', 
             'EU27_2020 - European Union - 27 countries (from 2020)'), 
         year_month >= 'Jan 2010') %>% 
  ggplot(mapping = aes(x = year_month, y = index_2015_100)) + 
  geom_line(data = . %>% filter(category == 'All-items HICP'), 
            #aes(group = iso3),
            size = .8, color = "gray80") + 
  geom_line(data= . %>% filter(category == 'Books'),
            color = "firebrick",
            lineend = "round") + 
  geom_point(data = . %>% filter(year_month == 'Aug 2020',
                                 category == 'Books'), 
             size = rel(1), 
             shape = 21, 
             color = "firebrick",
             fill = "firebrick2") + 
  geom_text(data= country_panel,
            mapping = aes(label = country,
                          x = year_month,
                          y = 185),
            vjust = "inward",
            hjust = "inward",
            fontface = "bold",
            color = "firebrick") +
  facet_wrap(~ country, ncol = 5) + 
  #facet_wrap(~ reorder(cname, -cu_cases), ncol = 5) + 
  labs(x='',
       y='CPI, 2015 = 100\n',
       color='', 
       caption = 'Makeover Monday, Week 38') + 
  theme_minimal() + 
  theme(plot.title = element_text(size = rel(1), face = "bold"),
        plot.subtitle = element_text(size = rel(0.7)),
        plot.caption = element_text(size = rel(1)),
        strip.text = element_blank(),
        panel.spacing.x = unit(-0.05, "lines"),
        panel.spacing.y = unit(0.3, "lines"),
        #axis.text.y = element_text(size = rel(0.5)),
        axis.title.x = element_text(size = rel(1)),
        axis.title.y = element_text(size = rel(1)),
        axis.text.x = element_text(size = rel(.75), angle = 90),
        legend.text = element_text(size = rel(1)))
```

## Mapping  

Attempting to highlight story with a map. More analysis would need to be done to determine whether it is meaningful.

```{r df_2, include=TRUE}
df_2 <- df %>% 
  spread(category, index_2015_100) %>% 
  mutate(cost_diff = Books - `All-items HICP`)
```

```{r map_world, include=TRUE}
map_world <- map_data("world")

map_world %<>% 
  mutate(region = recode(region,
                         'Czech Republic' = 'Czechia',
                         'Macedonia' = 'North Macedonia',
                         'UK' = 'United Kingdom',
                         'USA' = 'United States'))
```

```{r join_map, include=TRUE}
df_2 %<>% 
  left_join(x=.,
            y=map_world,
            by=c('country' = 'region'))
```

```{r map_diff, fig.height=8}
df_2 %>% 
  filter(year_month == 'Aug 2020',
         country != 'United States') %>% 
  ggplot(.) +
  geom_polygon(aes(x = long, y = lat, group = group, fill = cost_diff), color="gray60") +
  scale_fill_viridis() + 
  #scale_fill_gradient2() + 
  #scale_fill_brewer(palette = 'RdBu')
  ggtitle('Where are books more or less expensive?',
          subtitle = 'Cost differential between the Harmonized Index of Consumer Prices (HCPI) and the cost of books') + 
  labs(caption = 'August 2020') + 
  map.theme + 
  theme(plot.title = element_text(size=24),
        plot.subtitle = element_text(size=18))
```
